---
title: "SocialMobility_sensitivityanalyses_01182022"
author: "Gloria Huei-Jong Graf"
date: "1/18/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

This document contains reviewer-requested sensitivity analyses for Graf/SocialMobility/PNAS manuscript as of 01/18/2022.


```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(foreign)
library(modelr)
library(miceadds)
library(haven)
library(wesanderson)
library(patchwork)
library(lattice)
library(gridExtra)
library(corrplot)
```

## Loading data
```{r echo=FALSE}
# Main Dataset
socialmobility_df = read.dta("./Data/HRS_SocialMobilityStata12_DWB200823.dta") %>%
    janitor::clean_names() %>%
    arrange(hhidpn)
    #n=37722

# 2018 tracker file
trackerfile_df = read_sas("./Data/trk2018tr_r.sas7bdat") %>%
  janitor::clean_names() %>%
  mutate(hhidpn = paste(hhid, pn, sep = "")) %>%
  mutate(hhidpn = as.numeric(hhidpn)) %>%
  select(hhidpn, hhid, page, vbsi16wgtra, pvbswgtr, hispanic) %>%
  filter(page != "999") %>%
  arrange(hhidpn)
  #n=20911

# Vable childhood SES variables
vable_df = read_sas("./Data/cses_measures.sas7bdat") %>%
  janitor::clean_names() %>%
  mutate(hhidpn = paste(hhid, pn, sep = "")) %>%
  mutate(hhidpn = as.numeric(hhidpn)) %>%
  select(hhidpn, fincap, soccap, humcap, cses_index) %>%
  mutate(cses_index_pct = ((rank(cses_index)/length(cses_index))*100))

# Biological Age Datasets
## Phenoage
phenoage_df = read.dta("./Data/PhenoAge_Revised032320DK_v12.dta") %>%
    janitor::clean_names() %>%
    arrange(hhidpn) %>%
    select(hhidpn, pheno_age, paa) %>%
    filter(!is.na(pheno_age)) %>%
    rename(levinephenoage = pheno_age,
           levinephenoage_advance = paa)
  #n=9356
## KDM and HD
bioage_df =
  read_csv("./Data/new_bioage_df.csv") %>%
  rename(hhidpn = sampleID,
         ) %>%
  dplyr::filter(!is.na(hd))
  #n=9472

## Methylation clocks
methylclock_df = read.dta("./Data/EPICLOCKA_R.dta") %>%
    janitor::clean_names() %>%
    arrange(hhidpn) %>%
    dplyr::select(hhidpn, levine_dnamage, hannum_dnamage, horvath_dnamage, dnamgrimage, mpoa) %>%
    rename(levinednam = levine_dnamage,
           hannum = hannum_dnamage, 
           horvath = horvath_dnamage, 
           grimage = dnamgrimage,
           poa = mpoa) %>%
    dplyr::mutate(hhidpn = as.integer(hhidpn))
#n=4018
```

Data for adult wealth: 

1) We use data from all waves at which the respondent indicated that they were not retired (RwRETEMP = 0). We then then inflate to constant dollars, and calculate their max/min income and earnings during this time (RwIWAGE, RwIEARN). We repeat the process for the respondent's spouse (SwIWAGE, SwIEARN). We then set the respondents value to the higher of their/their spouse's income and earnings based on the higher-earning spouse.

```{r warning=FALSE, echo=FALSE}
#Loading datasets
RAND_df = read.dta("./Data/randhrs1992_2018v1.dta") %>%
  janitor::clean_names() %>%
  mutate(hhidpn = paste(hhid, pn, sep = "")) %>%
  mutate(hhidpn = as.numeric(hhidpn)) %>%
  arrange(hhidpn)

RANDimpute_df = read_sas("./Data/randhrsimp1992_2018v1.sas7bdat") %>%
  janitor::clean_names() %>%
  mutate(hhidpn = paste(hhid, pn, sep = "")) %>%
  mutate(hhidpn = as.numeric(hhidpn)) %>%
  arrange(hhidpn)

#Restricting variables
earnings_df =
  RAND_df %>%
  select(hhidpn,
         #respondent retirement status
         r1retemp, r2retemp, r3retemp, r4retemp, r5retemp, r6retemp, r7retemp, r8retemp, r9retemp, r10retemp, r11retemp, r12retemp, r13retemp, r14retemp,
         #spouse retirement status
         s1retemp, s2retemp, s3retemp, s4retemp, s5retemp, s6retemp, s7retemp, s8retemp, s9retemp, s10retemp, s11retemp, s12retemp, s13retemp, s14retemp,
         #respondent earnings
         r1iearn, r2iearn, r3iearn, r4iearn, r5iearn, r6iearn, r7iearn, r8iearn, r9iearn, r10iearn, r11iearn, r12iearn, r13iearn, r14iearn,
         #spousal earnings
         s1iearn, s2iearn, s3iearn, s4iearn, s5iearn, s6iearn, s7iearn, s8iearn, s9iearn, s10iearn, s11iearn, s12iearn, s13iearn, s14iearn) %>%
  mutate_at(vars(r1iearn:r14iearn), .funs = funs(. / 1000)) %>%
  mutate_at(vars(s1iearn:s14iearn), .funs = funs(. / 1000))

wages_df =
  RANDimpute_df %>%
  select(hhidpn,
        #respondent wages
         r1iwage, r2iwage, r3iwage, r4iwage, r5iwage, r6iwage, r7iwage, r8iwage, r9iwage, r10iwage, r11iwage, r12iwage, r13iwage, r14iwage,
         #spousal wages
         s1iwage, s2iwage, s3iwage, s4iwage, s5iwage, s6iwage, s7iwage, s8iwage, s9iwage, s10iwage, s11iwage, s12iwage, s13iwage, s14iwage) %>%
  mutate_at(vars(r1iwage:r14iwage), .funs = funs(. / 1000)) %>%
  mutate_at(vars(s1iwage:s14iwage), .funs = funs(. / 1000))
    
#Merge wages and earnings data, create non-retirement income variable
attainments_df =
  left_join(wages_df, earnings_df) %>%
  filter(r1retemp == "0.no retire empstat" | r2retemp == "0.no retire empstat" | r3retemp == "0.no retire empstat" | r4retemp == "0.no retire empstat" | r5retemp == "0.no retire empstat" | r6retemp == "0.no retire empstat" | r7retemp == "0.no retire empstat" | r8retemp == "0.no retire empstat" | r9retemp == "0.no retire empstat" | r10retemp == "0.no retire empstat" | r11retemp == "0.no retire empstat" | r12retemp == "0.no retire empstat" | r13retemp == "0.no retire empstat" | r14retemp == "0.no retire empstat")

#Write function
nonret_var = function(ownret, ownvar, spouseret, spousevar) {
    
  own_attainment_nonret = ifelse(pull(attainments_df, ownret) == "0.no retire empstat", pull(attainments_df, ownvar), NA)
  spouse_attainment_nonret = ifelse(pull(attainments_df, spouseret) == "0.no retire empstat", pull(attainments_df, spousevar), NA)
  hhidpn = attainments_df$hhidpn
  
  tibble(own_attainment_nonret = own_attainment_nonret,
         spouse_attainment_nonret = spouse_attainment_nonret,
         wave = paste(ownvar),
         hhidpn = hhidpn)
  
  }

### Analysis combinations
sample = list(attainments_df)

ownret = c("r1retemp", "r2retemp", "r3retemp", "r4retemp", "r5retemp", "r6retemp", "r7retemp", "r8retemp", "r9retemp", "r10retemp", "r11retemp", "r12retemp", "r13retemp", "r14retemp")

spouseret = c("s1retemp", "s2retemp", "s3retemp", "s4retemp", "s5retemp", "s6retemp", "s7retemp", "s8retemp", "s9retemp", "s10retemp", "s11retemp", "s12retemp", "s13retemp", "s14retemp")

ownearn = c("r1iearn", "r2iearn", "r3iearn", "r4iearn", "r5iearn", "r6iearn", "r7iearn", "r8iearn", "r9iearn", "r10iearn", "r11iearn", "r12iearn", "r13iearn", "r14iearn")

spouseearn = c("s1iearn", "s2iearn", "s3iearn", "s4iearn", "s5iearn", "s6iearn", "s7iearn", "s8iearn", "s9iearn", "s10iearn", "s11iearn", "s12iearn", "s13iearn", "s14iearn")

ownwage = c("r1iwage", "r2iwage", "r3iwage", "r4iwage", "r5iwage", "r6iwage", "r7iwage", "r8iwage", "r9iwage", "r10iwage", "r11iwage", "r12iwage", "r13iwage", "r14iwage")

spousewage = c("s1iwage", "s2iwage", "s3iwage", "s4iwage", "s5iwage", "s6iwage", "s7iwage", "s8iwage", "s9iwage", "s10iwage", "s11iwage", "s12iwage", "s13iwage", "s14iwage")


# Pre-retirement earnings calculations
earnings_preret = 
  data.frame(ownret, ownearn, spouseret, spouseearn)

earnings_preret_df =
  pmap_dfr(list(earnings_preret$ownret, earnings_preret$ownearn, earnings_preret$spouseret, earnings_preret$spouseearn), nonret_var) %>%
  mutate(wave = str_replace(wave, "iearn", "")) %>%
  mutate(wave = str_replace(wave, "r", "")) %>%
  rename(rearn_nonret_raw = own_attainment_nonret,
         searn_nonret_raw = spouse_attainment_nonret) %>%
  mutate(cpi_index = case_when(wave == 1 ~ 210.2,
                               wave == 2 ~ 220.0,
                               wave == 3 ~ 231.3,
                               wave == 4 ~ 239.5,
                               wave == 5 ~ 252.9,
                               wave == 6 ~ 264.2,
                               wave == 7 ~ 277.5,
                               wave == 8 ~ 296.2,
                               wave == 9 ~ 316.3,
                               wave == 10 ~ 320.4,
                               wave == 11 ~ 337.5,
                               wave == 12 ~ 348.3,
                               wave == 13 ~ 353.4,
                               wave == 14 ~ 369.8,
                               )) %>%
  mutate(rearn_nonret_adj = rearn_nonret_raw*(cpi_index/337.5),
          searn_nonret_adj = searn_nonret_raw*(cpi_index/337.5)) %>%
  pivot_wider(names_from = wave, values_from = c(cpi_index, rearn_nonret_raw, searn_nonret_raw, rearn_nonret_adj, searn_nonret_adj))

    # Mean and max calculations (NOT YET FINAL- 7408 with 0 income? out of 34309?)
    earnings_preret_df =
      earnings_preret_df %>%
      mutate(ownearn_mean = rowMeans(select(., starts_with("rearn_nonret_")), na.rm = TRUE),
             spouseearn_mean = rowMeans(select(., starts_with("searn_nonret_")), na.rm = TRUE)
             ) %>%
      mutate(ownearn_mean = ifelse(is.nan(ownearn_mean), NA, ownearn_mean),
             spouseearn_mean = ifelse(is.nan(spouseearn_mean), NA, spouseearn_mean)
             ) %>%
      rowwise %>%
      mutate(ownearn_max = max(c_across(rearn_nonret_adj_1:rearn_nonret_adj_14), na.rm = TRUE),
             spouseearn_max = max(c_across(searn_nonret_adj_1:searn_nonret_adj_14), na.rm = TRUE)
             ) %>%
      ungroup() %>%
      mutate(ownearn_max = ifelse(is.infinite(ownearn_max), NA, ownearn_max),
             spouseearn_max = ifelse(is.infinite(spouseearn_max), NA, spouseearn_max)
             ) %>%
      rowwise() %>%
      mutate(maxHH_meanearn = max(c_across(ownearn_mean:spouseearn_mean), na.rm = T),
             maxHH_maxearn = max(c_across(ownearn_max:spouseearn_max), na.rm = T)
               ) %>%
      ungroup() %>%
      mutate(maxHH_meanearn = ifelse(is.infinite(maxHH_meanearn), NA, maxHH_meanearn),
             maxHH_maxearn = ifelse(is.infinite(maxHH_maxearn), NA, maxHH_maxearn)
             ) %>%
      filter(!is.na(maxHH_meanearn)) %>%
      mutate(pct_maxHH_meanearn = ((rank(maxHH_meanearn)/length(maxHH_meanearn))*100),
             z_maxHH_meanearn = scale(maxHH_meanearn),
             pct_maxHH_maxearn = ((rank(maxHH_maxearn)/length(maxHH_maxearn))*100),
             z_maxHH_maxearn = scale(maxHH_maxearn)
             ) %>%
      filter(maxHH_meanearn != 0)
    

# Pre-retirement wage calculations
wages_preret = 
  data.frame(ownret, ownwage, spouseret, spousewage)

wages_preret_df =
  pmap_dfr(list(wages_preret$ownret, wages_preret$ownwage, wages_preret$spouseret, wages_preret$spousewage), nonret_var) %>%
  mutate(wave = str_replace(wave, "iwage", "")) %>%
  mutate(wave = str_replace(wave, "r", "")) %>%
  rename(rwage_nonret_raw = own_attainment_nonret,
         swage_nonret_raw = spouse_attainment_nonret) %>%
  mutate(cpi_index = case_when(wave == 1 ~ 210.2,
                               wave == 2 ~ 220.0,
                               wave == 3 ~ 231.3,
                               wave == 4 ~ 239.5,
                               wave == 5 ~ 252.9,
                               wave == 6 ~ 264.2,
                               wave == 7 ~ 277.5,
                               wave == 8 ~ 296.2,
                               wave == 9 ~ 316.3,
                               wave == 10 ~ 320.4,
                               wave == 11 ~ 337.5,
                               wave == 12 ~ 348.3,
                               wave == 13 ~ 353.4,
                               wave == 14 ~ 369.8,
                               )) %>%
  mutate(rwage_nonret_adj = rwage_nonret_raw*(cpi_index/337.5),
          swage_nonret_adj = swage_nonret_raw*(cpi_index/337.5)) %>%
  pivot_wider(names_from = wave, values_from = c(cpi_index, rwage_nonret_raw, swage_nonret_raw, rwage_nonret_adj, swage_nonret_adj))

    # Mean and max calculations (NOT YET FINAL)
    wages_preret_df =
      wages_preret_df %>%
      mutate(ownwage_mean = rowMeans(select(., starts_with("rwage_nonret_")), na.rm = TRUE),
             spousewage_mean = rowMeans(select(., starts_with("swage_nonret_")), na.rm = TRUE)) %>%
      mutate(ownwage_mean = ifelse(is.nan(ownwage_mean), NA, ownwage_mean),
             spousewage_mean = ifelse(is.nan(spousewage_mean), NA, spousewage_mean)) %>%
      rowwise() %>%
      mutate(ownwage_max = max(c_across(rwage_nonret_adj_1:rwage_nonret_adj_14), na.rm = TRUE),
             spousewage_max = max(c_across(swage_nonret_adj_1:swage_nonret_adj_14), na.rm = TRUE)
             ) %>%
      ungroup() %>%
      mutate(ownwage_max = ifelse(is.infinite(ownwage_max), NA, ownwage_max),
             spousewage_max = ifelse(is.infinite(spousewage_max), NA, spousewage_max)
             ) %>%
      rowwise() %>%
      mutate(maxHH_meanwage = max(c_across(ownwage_mean:spousewage_mean), na.rm = T),
             maxHH_maxwage = max(c_across(ownwage_max:spousewage_max), na.rm = T)
               ) %>%
      ungroup() %>%
      mutate(maxHH_meanwage = ifelse(is.infinite(maxHH_meanwage), NA, maxHH_meanwage),
             maxHH_maxwage = ifelse(is.infinite(maxHH_maxwage), NA, maxHH_maxwage)
             ) %>%
      filter(!is.na(maxHH_meanwage)) %>%
      mutate(pct_maxHH_meanwage = ((rank(maxHH_meanwage)/length(maxHH_meanwage))*100),
             z_maxHH_meanwage = scale(maxHH_meanwage),
             pct_maxHH_maxwage = ((rank(maxHH_maxwage)/length(maxHH_maxwage))*100),
             z_maxHH_maxwage = scale(maxHH_maxwage)
             ) %>%
      filter(maxHH_meanwage != 0)
    
```

``` {r}
# Merge datasets
main_df = left_join(socialmobility_df, trackerfile_df, by = "hhidpn")
main_df = left_join(main_df, vable_df, by = "hhidpn")
main_df = left_join(main_df, phenoage_df, by = "hhidpn")
main_df = left_join(main_df, bioage_df, by = "hhidpn")
main_df = left_join(main_df, methylclock_df, by = "hhidpn")
main_df = left_join(main_df, earnings_preret_df, by = "hhidpn")
main_df = left_join(main_df, wages_preret_df, by = "hhidpn")
main_df =
  main_df %>%
  filter(!is.na(page)) %>%
  rename(age = page)

# Generate methylation clock advancement measures
levinednam_advance_reg =
  lm(levinednam ~ age, data = main_df)

main_df =
  main_df %>%
  add_predictions(levinednam_advance_reg, var = "levinednam_pred") %>%
  dplyr::mutate(levinednam_advance = levinednam - levinednam_pred)

hannum_advance_reg =
  lm(hannum ~ age, data = main_df)

main_df =
  main_df %>%
  add_predictions(hannum_advance_reg, var = "hannum_pred") %>%
  dplyr::mutate(hannum_advance = hannum - hannum_pred)

horvath_advance_reg =
  lm(horvath ~ age, data = main_df)

main_df =
  main_df %>%
  add_predictions(horvath_advance_reg, var = "horvath_pred") %>%
  dplyr::mutate(horvath_advance = horvath - horvath_pred)

grimage_advance_reg =
  lm(grimage ~ age, data = main_df)

main_df =
  main_df %>%
  add_predictions(grimage_advance_reg, var = "grimage_pred") %>%
  dplyr::mutate(grimage_advance = grimage - grimage_pred)

# Standardize biological and methylation aging clock advancement variables
main_df =
  main_df %>%
  dplyr::mutate(
         levinephenoageadv_sd = (levinephenoage_advance - mean(main_df$levinephenoage_advance, na.rm = T))/sd(main_df$levinephenoage_advance, na.rm = T),
         kdma_sd = (kdm_advance - mean(main_df$kdm_advance, na.rm = T))/sd(main_df$kdm_advance, na.rm = T),
         hdlog_sd = (hd_log - mean(main_df$hd_log, na.rm = T))/sd(main_df$hd_log, na.rm = T),
         levinednamadv_sd = (levinednam_advance - mean(main_df$levinednam_advance, na.rm = T))/sd(main_df$levinednam_advance, na.rm = T),
         hannumadv_sd = (hannum_advance - mean(main_df$hannum_advance, na.rm = T))/sd(main_df$hannum_advance, na.rm = T),
         horvathadv_sd = (horvath_advance - mean(main_df$horvath_advance, na.rm = T))/sd(main_df$horvath_advance, na.rm = T),
         grimageadv_sd = (grimage_advance - mean(main_df$grimage_advance, na.rm = T))/sd(main_df$grimage_advance, na.rm = T),
         poa_sd = (poa - mean(main_df$poa, na.rm = T))/sd(main_df$poa, na.rm = T)
         )

# Create mobility variables - delta
  # Original variable: pctso_bc (origins), pctwlth13 (attainments), dpct13_bc (mobility)
  # New variables: cses_index_pct (origins), pct_maxHH_meanwage (attainments)
  # Mobility variables: dpct13_bc (original), deltamobility_cses_wlth, deltamobility_so_meanwage, deltamobility_cses_meanwage

main_df = 
  main_df %>%
  mutate(deltamobility_cses_wlth = pctwlth13 - cses_index_pct,
         deltamobility_so_meanwage = pct_maxHH_meanwage - pctso_bc,
         deltamobility_cses_meanwage = pct_maxHH_meanwage - cses_index_pct
         )

# Create mobility variables - residualized-change
  # Original variable: pctso_bc (origins), pctwlth13 (attainments), dpct13_bc (mobility)
  # New variables: cses_index_pct (origins), pct_maxHH_meanwage (attainments)
  # Mobility variables: rpct13_bc (original), deltamobility_cses_wlth, deltamobility_so_meanwage, deltamobility_cses_meanwage

rcmobility_original_reg =
  lm(pctwlth13 ~ pctso_bc, data = main_df)
  
rcmobility_cses_wlth_reg =
  lm(pctwlth13 ~ cses_index_pct, data = main_df)

rcmobility_so_meanwage_reg = 
  lm(pct_maxHH_meanwage ~ pctso_bc, data = main_df)

rcmobility_cses_meanwage_reg = 
  lm(pct_maxHH_meanwage ~ cses_index_pct, data = main_df)

main_df =
  main_df %>%
  add_predictions(rcmobility_original_reg, var = "rcmobility_original_reg_pred") %>%
  add_predictions(rcmobility_cses_wlth_reg, var = "rcmobility_cses_wlth_reg_pred") %>%
  add_predictions(rcmobility_so_meanwage_reg, var = "rcmobility_so_meanwage_reg_pred") %>%
  add_predictions(rcmobility_cses_meanwage_reg, var = "rcmobility_cses_meanwage_reg_pred") %>%
  dplyr::mutate(
    rcmobility_original = pctwlth13 - rcmobility_original_reg_pred,
    rcmobility_cses_wlth = pctwlth13 - rcmobility_cses_wlth_reg_pred,
    rcmobility_so_meanwage = pct_maxHH_meanwage - rcmobility_so_meanwage_reg_pred,
    rcmobility_cses_meanwage = pct_maxHH_meanwage - rcmobility_cses_meanwage_reg_pred
    )


#Summary
summary(main_df$pctso_bc)
summary(main_df$pctwlth13)
summary(main_df$cses_index_pct)
summary(main_df$pct_maxHH_meanwage)

summary(main_df$dpct13_bc)
summary(main_df$deltamobility_cses_wlth)
summary(main_df$deltamobility_so_meanwage)
summary(main_df$deltamobility_cses_meanwage)

summary(main_df$rpct13_bc)
summary(main_df$rcmobility_cses_wlth)
summary(main_df$rcmobility_so_meanwage)
summary(main_df$rcmobility_cses_meanwage)

# Cleaning dataset
main_df =
  main_df %>%
  dplyr::select(
    hhid, hhidpn:racohbyr, hispanic, rafeduc, rameduc, age, zwlth13, pctwlth13, z_maxHH_maxearn, zso_bc, pctso_bc, pct_maxHH_meanwage, pct_maxHH_meanearn, pct_maxHH_maxwage, z_maxHH_meanwage, pct_maxHH_maxearn, z_maxHH_meanearn, fameduc, cses_index_pct, dpct13_bc, deltamobility_cses_wlth, deltamobility_so_meanwage, deltamobility_cses_meanwage, rpct13_bc, rcmobility_original, rcmobility_cses_wlth, rcmobility_so_meanwage, rcmobility_cses_meanwage, levinephenoage:poa_sd) %>%
  mutate(
    educ = ifelse(raeduc == "1.lt high-school" | raeduc == "2.ged", "<HS",
            ifelse(raeduc == "3.high-school graduate" | raeduc == "4.some college", "HS",
            ifelse(raeduc == "5.college and above", "BA+",
            ifelse(is.na(raeduc), NA, 666)))),
    paryrseduc = ifelse(is.na(rameduc) | ((rafeduc >= rameduc) & !is.na(rafeduc)) , rafeduc, rameduc)
  ) %>%
  mutate(educ = factor(educ, levels = c("<HS", "HS", "BA+")),
         pareduc = ifelse(paryrseduc < 12, "<HS",
                    ifelse(paryrseduc >= 12 & paryrseduc <= 15, "HS",
                    ifelse(paryrseduc >= 16, "BA+", 666))),
         agesquared = age^2,
         hispanic = ifelse(hispanic == "5", 0, 1),
         pctso_bc = pctso_bc/25,
         cses_index_pct = cses_index_pct/25,
         pctwlth13 = pctwlth13/25,
         pct_maxHH_meanwage = pct_maxHH_meanwage/25,
         pct_maxHH_meanearn = pct_maxHH_meanearn/25,
         pct_maxHH_maxearn = pct_maxHH_maxearn/25,
         dpct13_bc = dpct13_bc/25,
         deltamobility_cses_wlth = deltamobility_cses_wlth/25,
         deltamobility_so_meanwage = deltamobility_so_meanwage/25,
         deltamobility_cses_meanwage = deltamobility_cses_meanwage/25,
         rpct13_bc = rpct13_bc/25,
         rcmobility_original = rcmobility_original/25,
         rcmobility_cses_wlth = rcmobility_cses_wlth/25,
         rcmobility_so_meanwage = rcmobility_so_meanwage/25,
         rcmobility_cses_meanwage = rcmobility_cses_meanwage/25,
         ) %>%
  mutate(pareduc = factor(pareduc, levels = c("<HS", "HS", "BA+")),
         educ_num = ifelse(educ == "<HS", 1,
                      ifelse(educ == "HS", 2,
                      ifelse(educ == "BA+", 3,
                      ifelse(is.na(educ), NA, 666)))),
         pareduc_num = ifelse(pareduc == "<HS", 1,
                      ifelse(pareduc == "HS", 2,
                      ifelse(pareduc == "BA+", 3,
                      ifelse(is.na(pareduc), NA, 666)))),
         fameduc_num = ifelse(fameduc == "Low", 1,
                      ifelse(fameduc == "Middle", 2,
                      ifelse(fameduc == "High", 3,
                      ifelse(is.na(fameduc), NA, 666))))
         ) %>%
  mutate(edmob_yrs = educ_num - pareduc_num,
         edmob_index = educ_num - fameduc_num
         )
#n=20607
```

## Analysis samples
```{r echo=FALSE}
# Full sample (n=20683)
full_sample = 
  main_df
#n=20683

full_sample_NH = 
  full_sample %>%
  filter(hispanic == 0)
#n=17267

# Biomarker sample
BA_subsample =
  main_df %>%
  filter(!is.na(levinephenoage) & !is.na(kdm) & !is.na(hd_log))
#9255

BA_subsample_NH = 
  BA_subsample %>%
  filter(hispanic == 0)
#7898

# DNAm sample
DNAm_subsample =
  main_df %>%
  filter(!is.na(horvath) | !is.na(hannum) | !is.na(levinednam) | !is.na(grimage) | !is.na(poa))
#n=3976

DNAm_subsample_NH = 
  DNAm_subsample %>%
  filter(hispanic == 0)
#n=3426

# Education subsamples 
BA_subsample_ED =
  BA_subsample %>%
  filter(!is.na(edmob_yrs))
#n=8759

DNAm_subsample_ED =
  DNAm_subsample %>%
  filter(!is.na(edmob_yrs))
#n=3774

## For interaction term analysis- datasets restricted to Black and white participants
BA_subsample_bw =
  BA_subsample_NH %>%
  filter(raracem %in% c("1.white/caucasian", "2.black/african american"))
#n=8451

BA_subsample_bw_ED =
  BA_subsample_NH %>%
  filter(raracem %in% c("1.white/caucasian", "2.black/african american")) %>%
  filter(!is.na(edmob_yrs))
#n=8451

DNAm_subsample_bw =
  DNAm_subsample_NH %>%
  filter(raracem %in% c("1.white/caucasian", "2.black/african american"))
#n=3661

DNAm_subsample_bw_ED =
  DNAm_subsample_NH %>%
  filter(raracem %in% c("1.white/caucasian", "2.black/african american")) %>%
  filter(!is.na(edmob_yrs))
#n=3661

## For analysis of middle 50% of social origins distribution (zso_bc)
summary(pull(BA_subsample, zso_bc))

BA_subsample_middle = 
  BA_subsample %>%
  filter(zso_bc > -0.6751444 & zso_bc < 0.6251319)
#n=4619

summary(pull(DNAm_subsample, zso_bc))

DNAm_subsample_middle = 
  DNAm_subsample %>%
  filter(zso_bc > -0.62451 & zso_bc < 0.67453)
#n=1971
```


## Social origins distributions for each sample, by race and sex
```{r echo=FALSE}
#Combined distributions (counts)
ggplot() +
  geom_histogram(data = full_sample, aes(x = zso_bc, y = ..count..), color = "grey65", fill = "grey65", alpha = 0.2, bins = 30) +
  geom_histogram(data = BA_subsample, aes(x = zso_bc, y = ..count..), color = "tomato", fill = "tomato", alpha = 0.2, bins = 30) +
  geom_histogram(data = DNAm_subsample, aes(x = zso_bc, y = ..count..), color = "dodgerblue1", fill = "dodgerblue1", alpha = 0.2, bins = 30) +
  theme_minimal() +
  labs(title = "Social origins distribution, full sample") +
  xlab("Social Origins (z-score)") +
  ylab(NULL)

ggplot() +
  geom_histogram(data = full_sample, aes(x = cses_index_pct, y = ..count..), color = "grey65", fill = "grey65", alpha = 0.2, bins = 30) +
  geom_histogram(data = BA_subsample, aes(x = cses_index_pct, y = ..count..), color = "tomato", fill = "tomato", alpha = 0.2, bins = 30) +
  geom_histogram(data = DNAm_subsample, aes(x = cses_index_pct, y = ..count..), color = "dodgerblue1", fill = "dodgerblue1", alpha = 0.2, bins = 30) +
  theme_minimal() +
  labs(title = "Social origins distribution, full sample") +
  xlab("Social Origins (Vable)") +
  ylab(NULL)

ggplot() +
  geom_histogram(data = full_sample %>% filter(!is.na(raracem)), aes(x = zso_bc, y = ..count..), color = "grey65", fill = "grey65", alpha = 0.2, bins = 30) +
  geom_histogram(data = BA_subsample %>% filter(!is.na(raracem)), aes(x = zso_bc, y = ..count..), color = "tomato", fill = "tomato", alpha = 0.2, bins = 30) +
  geom_histogram(data = DNAm_subsample %>% filter(!is.na(raracem)), aes(x = zso_bc, y = ..count..), color = "dodgerblue1", fill = "dodgerblue1", alpha = 0.2, bins = 30) +
  theme_minimal() +
  labs(title = "Social origins distribution, by race") +
  facet_grid(cols = vars(raracem)) +
  xlab("Social Origins (z-score)") +
  ylab(NULL)

ggplot() +
  geom_histogram(data = full_sample %>% filter(!is.na(raracem)), aes(x = cses_index_pct, y = ..count..), color = "grey65", fill = "grey65", alpha = 0.2, bins = 30) +
  geom_histogram(data = BA_subsample %>% filter(!is.na(raracem)), aes(x = cses_index_pct, y = ..count..), color = "tomato", fill = "tomato", alpha = 0.2, bins = 30) +
  geom_histogram(data = DNAm_subsample %>% filter(!is.na(raracem)), aes(x = cses_index_pct, y = ..count..), color = "dodgerblue1", fill = "dodgerblue1", alpha = 0.2, bins = 30) +
  theme_minimal() +
  labs(title = "Social origins distribution, by race") +
  facet_grid(cols = vars(raracem)) +
  xlab("Social Origins (Vable)") +
  ylab(NULL)

ggplot() +
  geom_histogram(data = full_sample, aes(x = zso_bc, y = ..count..), color = "grey65", fill = "grey65", alpha = 0.2, bins = 30) +
  geom_histogram(data = BA_subsample, aes(x = zso_bc, y = ..count..), color = "tomato", fill = "tomato", alpha = 0.2, bins = 30) +
  geom_histogram(data = DNAm_subsample, aes(x = zso_bc, y = ..count..), color = "dodgerblue1", fill = "dodgerblue1", alpha = 0.2, bins = 30) +
  theme_minimal() +
  labs(title = "Social origins distribution, by gender") +
  facet_grid(cols = vars(ragender)) +
  xlab("Social Origins (z-score)") +
  ylab(NULL)

ggplot() +
  geom_histogram(data = full_sample, aes(x = cses_index_pct, y = ..count..), color = "grey65", fill = "grey65", alpha = 0.2, bins = 30) +
  geom_histogram(data = BA_subsample, aes(x = cses_index_pct, y = ..count..), color = "tomato", fill = "tomato", alpha = 0.2, bins = 30) +
  geom_histogram(data = DNAm_subsample, aes(x = cses_index_pct, y = ..count..), color = "dodgerblue1", fill = "dodgerblue1", alpha = 0.2, bins = 30) +
  theme_minimal() +
  labs(title = "Social origins distribution, by gender") +
  facet_grid(cols = vars(ragender)) +
  xlab("Social Origins (Vable)") +
  ylab(NULL)
```


# New mobility analyses

Input lists
```{r echo=FALSE, results=FALSE}
# Creating input lists: samples (2: biomarker and DNAm), 4 measures of social mobility

## Sample
coremodel_samples = list(BA_subsample = BA_subsample, DNAm_subsample = DNAm_subsample)
         
# Outcome measures
sens_mobilitymeasures = list(pctso_bc = "pctso_bc", cses_index_pct = "cses_index_pct", pctwlth13 = "pctwlth13", pct_maxHH_meanwage = "pct_maxHH_meanwage", dpct13_bc = "dpct13_bc", deltamobility_cses_wlth = "deltamobility_cses_wlth", deltamobility_so_meanwage = "deltamobility_so_meanwage", deltamobility_cses_meanwage = "deltamobility_cses_meanwage", rpct13_bc = "rpct13_bc", rcmobility_original = "rcmobility_original", rcmobility_cses_wlth = "rcmobility_cses_wlth", rcmobility_so_meanwage = "rcmobility_so_meanwage", rcmobility_cses_meanwage = "rcmobility_cses_meanwage")

```

Analysis combinations
```{r echo=FALSE}
### SES combination
model1_combo_list_sens = 
  list(SESmeasure = sens_mobilitymeasures, df = coremodel_samples)
dataset_model1_combo_sens = 
  cross_df(model1_combo_list_sens)
```

```{r}
# Create mobility variables
  # Original variable: pctso_bc (origins), pctwlth13 (attainments), dpct13_bc (mobility)
  # New variables: cses_index_pct (origins), pct_maxHH_meanwage (attainments)
  # Mobility variables: dpct13_bc (original), deltamobility_cses_wlth, deltamobility_so_meanwage, deltamobility_cses_meanwage

mobilitymodel_f =
  function(SESmeasure, df) {

  lm.cluster(levinephenoageadv_sd ~
       pull(df, SESmeasure) + 
       age + agesquared + ragender + raracem + hispanic + age:ragender + agesquared:ragender, data = df, cluster = "hhid") %>%
  summary() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "term") %>%
  janitor::clean_names() %>%
  rename(std.error = std_error,
         p.value = pr_t) %>%
  filter(term == "pull(df, SESmeasure)") %>%
  mutate(lowerCI = format(round(estimate - 1.96*std.error, 2), nsmall = 2),
         upperCI = format(round(estimate + 1.96*std.error, 2), nsmall = 2)
         ) %>%
  mutate(CI = paste("[", lowerCI, ",", upperCI, "]", sep = ""),
         estimate = format(estimate, nsmall = 2),
          p.value = formatC(p.value, format = "e")
         ) %>%
  select(term, estimate, CI, p.value)

  }

nobs_mobilitymodel_f =
  function(SESmeasure, df) {

  model = lm(levinephenoageadv_sd ~
       pull(df, SESmeasure) + 
       age + agesquared + ragender + raracem + hispanic + age:ragender + agesquared:ragender, data = df) 
  
  nobs(model) %>%
    as.data.frame()

  }

```

Sensitivity analysis: Mobility measures
```{r echo=FALSE, include=FALSE}
### Applying function
mobilitymodel_estimates = 
  pmap_dfr(list(dataset_model1_combo_sens$SESmeasure, dataset_model1_combo_sens$df), mobilitymodel_f) %>%
  mutate(
    SES_measure = rep(c("pctso_bc", "cses_index_pct", "pctwlth13", "pct_maxHH_meanwage", "dpct13_bc", "deltamobility_cses_wlth", "deltamobility_so_meanwage", "deltamobility_cses_meanwage", "rpct13_bc", "rcmobility_original", "rcmobility_cses_wlth", "rcmobility_so_meanwage", "rcmobility_cses_meanwage"), 2),
    data = rep(c("BA_subsample", "DNAm_subsample"), each = 13)
    ) %>%
  select(SES_measure, data, estimate, CI, p.value) %>%
  pivot_wider(names_from = data, values_from = c(estimate, CI, p.value)) %>%
  select(SES_measure, 
         estimate_BA_subsample, CI_BA_subsample, p.value_BA_subsample,
         estimate_DNAm_subsample, CI_DNAm_subsample, p.value_DNAm_subsample
         )

###nobs
mobilitymodel_estimates_nobs =
  pmap_dfr(list(dataset_model1_combo_sens$SESmeasure, dataset_model1_combo_sens$df), nobs_mobilitymodel_f) %>%
  mutate(
    SES_measure = rep(c("pctso_bc", "cses_index_pct", "pctwlth13", "pct_maxHH_meanwage", "dpct13_bc", "deltamobility_cses_wlth", "deltamobility_so_meanwage", "deltamobility_cses_meanwage", "rpct13_bc", "rcmobility_original", "rcmobility_cses_wlth", "rcmobility_so_meanwage", "rcmobility_cses_meanwage"), 2),
    data = rep(c("BA_subsample", "DNAm_subsample"), each = 13)
    ) %>%
  rename(n = names(.)[1]) %>%
  pivot_wider(names_from = data, values_from = n)


```

```{r echo=FALSE}
mobilitymodel_estimates %>%
  knitr::kable()

mobilitymodel_estimates_nobs %>%
  knitr::kable()

mobilitycor_BA =
  BA_subsample %>%
  select(dpct13_bc, deltamobility_cses_wlth, deltamobility_so_meanwage, deltamobility_cses_meanwage, rpct13_bc, rcmobility_original, rcmobility_cses_wlth, rcmobility_so_meanwage, rcmobility_cses_meanwage)

corrplot(corr=cor(mobilitycor_BA, use="complete.obs"), method="color", 
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
         )

```


## Supplemental analysis: effect of social origins controlling for attainment

Input lists
```{r echo=FALSE, results=FALSE}
## Sample
suppmodel_samples = list(BA_subsample = BA_subsample, DNAm_subsample = DNAm_subsample)

## BA measures
suppmodel_BAmeasures = list(levinephenoageadv_sd = "levinephenoageadv_sd", kdma_sd = "kdma_sd", hdlog_sd = "hdlog_sd", horvathadv_sd = "horvathadv_sd", hannumadv_sd = "hannumadv_sd", levinednamadv_sd = "levinednamadv_sd", grimageadv_sd = "grimageadv_sd", poa_sd = "poa_sd")

# Outcome measures
suppmodel_originmeasures = list(pctso_bc = "pctso_bc", cses_index_pct = "cses_index_pct")
suppmodel_attainmentmeasures = list(pctwlth13 = "pctwlth13", pct_maxHH_meanwage = "pct_maxHH_meanwage")

```

Analysis combinations
```{r echo=FALSE}

### SES combination
suppmodel_combo_list = 
  list(BAmeasure = suppmodel_BAmeasures, originsmeasure = suppmodel_originmeasures, attainmentsmeasure = suppmodel_attainmentmeasures, df = coremodel_samples)
dataset_suppmodel_combo_list = 
  cross_df(suppmodel_combo_list)

```

Function writing
```{r echo=FALSE, results=FALSE}
# # function foundation example
# test =
#   lm.cluster(levinephenoageadv_sd ~ pctso_bc + age + agesquared + ragender + raracem + hispanic + age:ragender + agesquared:ragender, data = DNAm_subsample, cluster = "hhid") %>%
#   summary() %>%
#   as.data.frame() %>%
#   tibble::rownames_to_column(var = "term") %>%
#   janitor::clean_names() %>%
#   rename(std.error = std_error,
#          p.value = pr_t) %>%
#   filter(term == "pctso_bc") %>%
#   mutate(lowerCI = round(estimate - 1.96*std.error, 2),
#          upperCI = round(estimate + 1.96*std.error, 2)
#          ) %>%
#   mutate(CI = paste("[", lowerCI, ",", upperCI, "]", sep = "")) %>%
#   select(term, estimate, CI, p.value)

### Function
suppmodel_f =
  function(BAmeasure, originsmeasure, attainmentsmeasure, df) {

  lm.cluster(pull(df, BAmeasure) ~ 
       pull(df, originsmeasure) + 
       age + agesquared + ragender + raracem + hispanic + age:ragender + agesquared:ragender + pull(df, attainmentsmeasure), data = df, cluster = "hhid") %>%
  summary() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "term") %>%
  janitor::clean_names() %>%
  rename(std.error = std_error,
         p.value = pr_t) %>%
  filter(term == "pull(df, originsmeasure)") %>%
  mutate(lowerCI = format(round(estimate - 1.96*std.error, 2), nsmall = 2),
         upperCI = format(round(estimate + 1.96*std.error, 2), nsmall = 2)
         ) %>%
  mutate(CI = paste("[", lowerCI, ",", upperCI, "]", sep = ""),
         estimate = format(estimate, nsmall = 2),
          p.value = formatC(p.value, format = "e")
         ) %>%
  select(term, estimate, CI, p.value)

  }

# use cases
# suppmodel_f(BAmeasure = "grimageadv_sd", "pctso_bc", "pctwlth13", df = DNAm_subsample)
# suppmodel_f(BAmeasure = "levinephenoageadv_sd", "cses_index_pct", "pct_maxHH_meanwage", df = BA_subsample)

```

Table 2- SES measures
```{r echo=FALSE, include=FALSE}

dataset_suppmodel_combo_list



### Applying function
suppmodel_SES_estimates = 
  pmap_dfr(list(dataset_suppmodel_combo_list$BAmeasure, dataset_suppmodel_combo_list$originsmeasure, dataset_suppmodel_combo_list$attainmentsmeasure, dataset_suppmodel_combo_list$df), suppmodel_f) %>%
  mutate(
    bioage_measure = rep(c("levinephenoageadv_sd", "kdma_sd", "hdlog_sd", "horvathadv_sd", "hannumadv_sd", "levinednamadv_sd", "grimageadv_sd", "poa_sd"), 8),
    originsmeasure = rep(c("pctso_bc", "cses_index_pct"), each = 8, 4),
    attainmentsmeasure = rep(c("pctwlth13", "pct_maxHH_meanwage"), each = 16, 2),
    data = rep(c("BA_subsample", "DNAm_subsample"), each = 32)
    ) %>%
  select(bioage_measure, originsmeasure, attainmentsmeasure, data, estimate, CI, p.value) %>%
  pivot_wider(names_from = data, values_from = c(estimate, CI, p.value)) %>%
  select(bioage_measure, originsmeasure, attainmentsmeasure, 
         estimate_BA_subsample, CI_BA_subsample, p.value_BA_subsample,
         estimate_DNAm_subsample, CI_DNAm_subsample, p.value_DNAm_subsample,
         )
```

```{r echo=FALSE}
suppmodel_SES_estimates %>%
  knitr::kable()
```



